Discriminating the Linear Influences of Collinear Factors \\from Multivariate Observations: the Back-to-Back Regression

### 2020-03-04 paper/appendix.tex
clean up

     Consider the B2B model from equation \ref{eq:model} $$Y = (XE + N)F$$ with $N$ centred and full rank noise.

@@

Samely, for $\hat H$, we have
\begin{align*}
    \hat H = \arg \min_H \mathbb{E}[\| XH - Y \hat{G} \|^2] &=\arg  \min_H \mathbb{E}[\| XH - (XE + N)F \hat G \|^2] \\
    &=\arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2] + \mathbb{E}[\| NF\hat G \| ^2]\\
    &= \arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2]
 \end{align*}
 a positive quantity which reaches a minimum (zero) for
 \begin{align}
    \hat H = EF \hat G
    \label{eq:Hdoublenom}
\end{align}

@@

Equation \ref{eq:model} does not explicitly contain a measurement noise term. Yet, in moast experimental cases, the problem is best described as:
\begin{equation}
  Y = (XE+N)F+M
  \label{eq:model_noisy_measure}
\end{equation}
  with $M\in \mathbb{R}^{n \times d_y}$.

@@

In case of noise, B2B yields non binary $\hat E$. Three thresholding rules can be used to binarize its values thus explicitly recover causal features.

First, given known signal-to-noise ratio,the threshold above which a feature should considered to be causal can be derived analytically. Indeed, Equation $\ref{eq:diagk}$ implies that the $k$ first diagonal elements of $\hat H$ are bounded:  % we assume that $F$ has full rank over $Img(E)$.

@@

The decision threshold between causal and non-causal elements is thus a fraction $\mu$, whose proportion arbitrarily depends on the necessity to favor type I and type II errors. In practice, we cannot use this procedure for our MEG study, because signal-to-noise ratio is unknown.

@@

This procedure assumes that there exists at least one causal and at least one non-causal feature.
%
Third, second-order statistics across multiple datasets can be used to identify the elements of $\text{diag}(\hat H)$ that are sgnificantly different from 0. This procedure is detailed in the method section of our MEG experiment.

Overall, these three procedure thus varies in their additional assumptions: i.e. (1) a known signal-to-noise ratio, (2) the existence of both causal and non-causal factors or (3) independent repetitions of the experiment.

>>>>

     Consider the B2B model from equation \ref{eq:model} $$Y = (XE + N)F$$ with
     $N$ centered and full rank noise.

@@

In the same way, for $\hat H$, we have
\begin{align*}
    \hat H = \arg \min_H \mathbb{E}[\| XH - Y \hat{G} \|^2] &=\arg  \min_H \mathbb{E}[\| XH - (XE + N)F \hat G \|^2] \\
    &=\arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2] + \mathbb{E}[\| NF\hat G \| ^2]\\
    &= \arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2]
 \end{align*}
 a positive quantity which reaches a minimum (zero) for
 \begin{align}
    \hat H = EF \hat G
    \label{eq:Hdoublenom}
\end{align}

@@

Equation \ref{eq:model} does not explicitly contain a measurement noise term.
Yet, in most experimental cases, the problem is best described as:
\begin{equation}
  Y = (XE+N)F+M
  \label{eq:model_noisy_measure}
\end{equation}
  with $M\in \mathbb{R}^{n \times d_y}$.

@@

In case of noise, B2B yields non binary $\hat E$. Three thresholding rules can
be used to binarize its values thus explicitly recover "causal" features.

First, given known signal-to-noise ratio, the threshold above which a feature
should considered to be "causal" can be derived analytically. Indeed, Equation
$\ref{eq:diagk}$ implies that the $k$ first diagonal elements of $\hat H$ are
bounded:  % we assume that $F$ has full rank over $Img(E)$.

@@

The decision threshold between "causal" and "non-causal" elements is thus a fraction
$\mu$, whose proportion arbitrarily depends on the necessity to favor type I and
type II errors. In practice, we cannot use this procedure for our MEG study,
because signal-to-noise ratio is unknown.

@@

This procedure assumes that there exists at least one causal and at least one
non-causal feature.
%
Third, second-order statistics across multiple datasets can be used to identify
the elements of $\text{diag}(\hat H)$ that are significantly different from 0.
This procedure is detailed in the method section of our MEG experiment.

Overall, these three procedures thus vary in their additional assumptions: i.e.
(1) a known signal-to-noise ratio, (2) the existence of both causal and
non-causal factors or (3) independent repetitions of the experiment.

### 2020-03-04 paper/introduction.tex
clean up

First, causal and non-causal factors may be numerous and collinear. %In physics, for example, one may be challenged to identify whether fusion is caused by a change in temperature or a change in pressure, as these two factors may, at first, be difficult to manipulate independently.
This issue becomes increasingly pronounced as the number of potential factors increases. In neuroscience, for example, identifying whether the frequency of a word presented on a subject's retina modulates brain activity can be surprisingly difficult. Indeed, the frequency of words in natural language covaries with other factors such as their length (short words are more frequent than long words), their categories (determinants are more frequent than adverbs) and so forth \citep{kutas2011thirty,pegado2014timing}. Instead of selecting a set of words that control for all of these factors simultaneously, it is thus common to use forward modeling, i.e. to train a model to predict observations (e.g. brain activity) from a minimal combination of competing factors (e.g. word length, word frequency, e.g. \citep{huth2016natural}), and investigate, in the model, the estimated contribution of each factor \citep{friston1994statistical}.

The second challenge to measuring causal influence is that observations can be large and complex. The relationship between causes and effects is thus often considered in a backward manner, by training models to maximally predict causes from multidimensional observations. For example, brain activity is often recorded with hundreds or thousands of sensors simultaneously. As multiple sensors may be affected by common noise sources, it is common to use backward regression, by, for example, fitting a support vector machine across multiple sensors to decode the category of a stimulus \citep{cichy2014resolving,  kriegeskorte2008representational, norman2006beyond}.

Both forward and backward modeling have competing benefits and drawbacks. Specifically, forward modeling disentangles the independent contribution of collinear factors, but does not combine multidimensional observations. By contrast, backward modeling combines multiple observations, but does not disentangle collinear factors \citep{weichwald2015causal, hebart2018deconstructing, king2018encoding}. To combine the benefits of forward and backward modeling, several authors have proposed to use cross-decomposition techniques such as Partial Least Squares (PLS) and Canonical Correlation Analysis (CCA) \citep{de2019multiway}.
CCA and PLS aim to find, from two sets of data $X$ and $Y$, the components $H$ and $G$ were $XH$ and $YG$ are maximially correlated or maximally covarying respectively. Because CCA and PLS are based on a generalized eigen decomposition, their resulting coefficients are mixing the features of $X$ and $Y$ in a way that makes them notoriously difficult to interpret \citep{lebart1995statistique}.

Here, we introduce the `back-to-back regression' (B2B), which not only combines the benefits of forward and backward modeling (Section~\ref{sec:algorithm}), but also provide robust, interpretable and unidimensional causal coefficients for each tested factor.

The present paper focuses on the restricted issue of disentangling the causal influence of linearly correlated predictors ($X$) onto noisy multivariate observations ($Y$). The present approach thus differs from other causal discovery algorithms based on temporal-delays and/or nonlinear interactions in systems where the directionality of causation (from X to Y or vice versa) is unknown (e.g. \citep{peters2017elements, granger1969investigating, janzing2013quantifying, scholkopf2016modeling}.

After detailing B2B method and proving its convergence (Section~\ref{sec:theorem}), we show with synthetic data that it outperforms state-of-the-art forward, backward and cross-decomposition techniques in identifying causal influence (Section~\ref{sec:experiment_synthetic}). Finally, we apply B2B to a large neuroimaging dataset and reveal that distinct but collinear word features lead to distinguishable brain representations (Section~\ref{sec:experiment_real}).

>>>>

First, causal and non-causal factors may be numerous and collinear.
% In physics,
% for example, one may be challenged to identify whether fusion is caused by a
% change in temperature or a change in pressure, as these two factors may, at
% first, be difficult to manipulate independently. This issue becomes increasingly
% pronounced as the number of potential factors increases.
In neuroscience, for
example, it can be challenging to identify whether word frequency modulates
brain activity during reading. Indeed, the
frequency of words in natural language covaries with other factors such as their
length (short words are more frequent than long words) and their categories
(determinants are more frequent than adverbs)
\citep{kutas2011thirty,pegado2014timing}. Instead of selecting a set of words
that controls for all of these factors simultaneously, it is thus common to use
a \emph{forward} "encoding model", i.e. to fit a linear regression to predict observations
(e.g. brain activity) from a minimal combination of competing factors (e.g.
word length, word frequency), and analytically investigate,
the estimated contribution of each factor from the model's coefficients
\citep{friston1994statistical,naselaris2011encoding,weichwald2015causal,
king2018encoding,huth2016natural}.

The second challenge to measuring causal influence is that observations can be
multidimensional. The relationship between causes and effects is thus often
considered in a \emph{backward} manner, by training models to maximally predict causes
from multidimensional observations. For example, brain activity is often
recorded with hundreds or thousands of simultaneous measurements via functional
Magnetic Resonance Imaging, magneto-encephalography (MEG) or multiple electro-
physiological probes \citep{friston1994statistical,steinmetz2018challenges}.
As simultaneous measurements may be affected by common noise sources, it is
common to use backward
modeling, by, for example, fitting a support vector machine across multiple
sensors to decode the category of a stimulus \citep{norman2006beyond,cichy2014resolving,
kriegeskorte2008representational}.

Both \emph{forward} and \emph{backward} modeling have competing benefits and drawbacks.
Specifically, forward modeling disentangles the independent contribution of
collinear factors, but does not combine multidimensional observations. By
contrast, backward modeling combines multiple observations, but does not
disentangle collinear factors \citep{weichwald2015causal,
hebart2018deconstructing, king2018encoding}. To combine some of the benefits of forward
and backward modeling, several authors have proposed to use cross-decomposition
techniques such as Partial Least Squares (PLS) and Canonical Correlation
Analysis (CCA) \citep{de2019multiway}. CCA and PLS aim to find, from two sets of
data $X$ and $Y$, the components $H$ and $G$ were $XH$ and $YG$ are maximally
correlated or maximally covarying respectively. Because CCA and PLS are based on
a generalized eigen decomposition, their resulting coefficients are mixing the
features of $X$ and $Y$ in a way that makes them notoriously difficult to
interpret \citep{lebart1995statistique}.

Here, we introduce the `back-to-back regression' (B2B), which not only combines
the benefits of forward and backward modeling (Section~\ref{sec:algorithm}), but
also provides robust, interpretable, unidimensional and unbiased coefficients for
each of tested factor.

The present paper focuses on the restricted issue of disentangling the
influence of linearly correlated predictors ($X$) onto noisy multivariate
observations ($Y$). The present approach thus differs from other causal
discovery algorithms based on temporal-delays and/or nonlinear interactions in
systems where the directionality of causation (from X to Y or vice versa) is
unknown (e.g. \citep{peters2017elements, granger1969investigating,
janzing2013quantifying, scholkopf2016modeling}.

After detailing B2B method and proving its convergence
(Section~\ref{sec:theorem}), we show with synthetic data that it outperforms
state-of-the-art forward, backward and cross-decomposition techniques in
disentangling causal factors (Section~\ref{sec:experiment_synthetic}). Finally,
we apply B2B to a large neuroimaging dataset and reveal that distinct but
collinear word features lead to distinguishable brain representations
(Section~\ref{sec:experiment_real}).

### 2020-03-04 paper/meg_exps.tex
clean up

The raw MEG data was bandpass-filtered between 0.1 and 40Hz using MNE-Python
default parameters \citep{gramfort2013meg, gramfort2014mne}. Specifically, we used a zero-phase finite impulse
response filter (FIR) with a Hamming window and with transition bands of 0.1Hz
and 10Hz for the low and high cut-off frequencies. The raw data was then segmented 100ms before word onset and 1s after
word onset ($t=0$ms corresponds to word onset). Finally, each resulting
segment was baseline-corrected between -100ms and 0ms, and decimated by 5 and
thus led a sampling frequency of 240Hz. The average responses across words is displayed in Figure \ref{fig:meg_evoked}.
For each subject and each time sample relative to word onset, we
build an observation matrix $Y \in \mathbb{R}^{n \times d_y}$ of $n\approx$ 2700 words
by $d_y=301$ MEG channels (273 magnetometers and 28 compensation channels). Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

\begin{figure}[t!]
  \begin{minipage}[c]{0.6\textwidth}
    \includegraphics[width=\textwidth, trim=0cm 0cm 0cm 0cm, clip=True]{figures/meg_evoked.pdf}
  \end{minipage}\hfill
  \begin{minipage}[c]{0.4\textwidth}
    \caption{Ninety subjects each read approximately 2,700 words while their brain activity was recorded with MEG. Top. Average brain response to words (word onset at t=0 ms), as viewed from above the head (red= higher gradient of magnetic flux). Bottom. Each line represents a magnetometer, color-coded by its spatial position. Posterior responses, typical of primary visual cortex activity, peak around 100 ms after word onset and are followed by an anterior propagation of activity typical of semantic processing in the associative cortices.
    }
    \label{fig:meg_evoked}
  \end{minipage}
\end{figure}

@@

We aim to identify the word features that cause a variation in brain responses. We consider four distinct but colinear features.
%
First, 'Word Length' refers to the total number of letters. Word Length is expected to specifically cause a variation in the early evoked MEG responses (i.e. from 100 ms after stimulus onset) elicited by the retinotopically-tuned visual cortices (e.g. \citep{pegado2014timing}.).
%
Second, 'Word Frequency' indexes how frequently each word appears in Dutch and was derived with the the Zipf logarithmic scale of \citep{van2014subtlex} provided by the WordFreq package \citep{speerwordfreq}. Word Frequency is expected to specifically cause a variation in the late evoked MEG responses (i.e. from 400 ms), because it variably engages semantic processes in the temporal cortices \citep{kutas2011thirty}.
%
Third, 'Word Function' indicates whether each word is a content word (i.e. a noun, a verb, an adjective or an adverb) or a function word (i.e. a preposition, a conjunction, a determinant, a pronoun or a numeral), and was derived from Spacy's part of speech tagger \citep{spacy2}. To our knowledge, this feature has not been thouroughly investigated with MEG. Its causal contribution to reading processes in the brain thus remains unclear.
%
Finally, to verify that B2B and other methods would not inadequately identify non-causal features, we added a dummy feature, constructed from a noisy combination of Word Length and Word Frequency:
$dummy = z(length) + z(frequency) + \mathcal{N}$, where $z$ normalizes features and $\mathcal{N}$ is a random vector sampling Gaussian distribution (all terms thus have a zero-mean and a unit-variance). This procedure yields an $X \in \mathbb{R}^{n \times d_x}$ matrix of $n\approx$ 2700 words by
$d_x=4$ features for each subject. Each of the columns of $X$ is normalized to
have a mean and a standard deviation of 0 and 1 respectively.

@@

We used the feature importance described in Algorithm \ref{algorithm:b2b_fi} to assess the extent to which each feature $X_i$ specifically improves the prediction of held-out $Y$ data, using a 5-fold cross-validation (with shuffled trials to homogeneize the distributions between the training and testing splits).

@@

As expected, the Backward model reveals a similar decoding time course for Word Length and Word Frequency, even though these features are known to specifically influence early and late MEG responses respectively \citep{kutas2011thirty}. In addition, the same decoding time course was observed for the dummy variable. These results illustrate that backward modelling cannot be used to estimate the causal contribution of collinear features.

@@

\begin{wrapfigure}{r}{.6\textwidth}
  \vspace{-12ex}
  \begin{center}
    \includegraphics[width=0.6\textwidth, trim=0cm 0cm 0cm 0cm, clip=True]{figures/meg.pdf}
  \end{center}
  \vspace{-5ex}
  \caption{Multiple models (color-coded) are compared on their ability to reliably predict single-trial MEG signals evoked by words. Left. Average improvement of correlation coefficient $\Delta R$ for each of the four features (rows). Error bars indicate standard error of the mean (SEM) across subjects. Right. Average $\Delta R$ across time for each subject (dots). Top horizontal lines indicate when B2B significantly outperforms other methods (red) and vice versa. \label{fig:meg_results}
}
\end{wrapfigure}

>>>>

The raw MEG data was bandpass-filtered between 0.1 and 40Hz using MNE-Python
default parameters \citep{gramfort2013meg, gramfort2014mne}. Specifically, we
used a zero-phase finite impulse
response filter (FIR) with a Hamming window and with transition bands of 0.1Hz
and 10Hz for the low and high cut-off frequencies. The raw data was then
segmented 100ms before word onset and 1s after
word onset ($t=0$ms corresponds to word onset). Finally, each resulting
segment was baseline-corrected between -100ms and 0ms, and decimated by 5 and
thus led a sampling frequency of 240Hz. The average responses across words is
displayed in Figure \ref{fig:meg_evoked}.
For each subject and each time sample relative to word onset, we
build an observation matrix $Y \in \mathbb{R}^{n \times d_y}$ of $n\approx$ 2,700 words
by $d_y=301$ MEG channels (273 magnetometers and 28 compensation channels). Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

\begin{figure}[t!]
  \begin{minipage}[c]{0.6\textwidth}
    \includegraphics[width=\textwidth, trim=0cm 0cm 0cm 0cm, clip=True]{figures/meg_evoked.pdf}
  \end{minipage}\hfill
  \begin{minipage}[c]{0.4\textwidth}
    \caption{A hundred subjects read approximately 2,700 words while their
    brain activity was recorded with MEG. Top. Average brain response to words
    (word onset at t=0 ms), as viewed from above the head (red= higher gradient
    of magnetic flux). Bottom. Each line represents a magnetometer, color-coded
    by its spatial position. Posterior responses, typical of primary visual
    cortex activity, peak around 100 ms after word onset and are followed by
    an anterior propagation of activity typical of semantic processing in the
    associative cortices.
    }
    \label{fig:meg_evoked}
  \end{minipage}
\end{figure}

@@

We aim to identify the word features that cause a variation in brain responses.
We consider four distinct but collinear features.
%
First, 'Word Length' refers to the total number of letters. Word Length is
expected to specifically cause a variation in the early evoked MEG responses
(i.e. from 100 ms after stimulus onset) elicited by the retinotopically-tuned
visual cortices (e.g. \citep{pegado2014timing}.).
%
Second, 'Word Frequency' indexes how frequently each word appears in Dutch and
was derived with the the Zipf logarithmic scale of \citep{van2014subtlex}
provided by the WordFreq package \citep{speerwordfreq}. Word Frequency is
expected to specifically cause a variation in the late evoked MEG responses
(i.e. from 400 ms), because it variably engages semantic processes in the
temporal cortices \citep{kutas2011thirty}.
%
Third, 'Word Function' indicates whether each word is a content word (i.e. a
noun, a verb, an adjective or an adverb) or a function word (i.e. a preposition,
a conjunction, a determinant, a pronoun or a numeral), and was derived from
Spacy's part of speech tagger \citep{spacy2}. To our knowledge, this feature has
not been thoroughly investigated with MEG. Its causal contribution to reading
processes in the brain thus remains unclear.
%
Finally, to verify that B2B and other methods would not inadequately identify
non-causal features, we added a dummy feature, constructed from a noisy
combination of Word Length and Word Frequency: $dummy = z(length) + z(frequency)
+ \mathcal{N}$, where $z$ normalizes features and $\mathcal{N}$ is a random
vector sampling Gaussian distribution (all terms thus have a zero-mean and a
unit-variance). This procedure yields an $X \in \mathbb{R}^{n \times d_x}$
matrix of $n\approx$ 2,700 words by $d_x=4$ features for each subject. Each of
the columns of $X$ is normalized to have a mean and a standard deviation of 0
and 1 respectively.

@@

We used the feature importance described in Algorithm \ref{algorithm:b2b_fi} to
assess the extent to which each feature $X_i$ specifically improves the
prediction of held-out $Y$ data, using a five-fold cross-validation (with shuffled
trials to homogeneize the distributions between the training and testing
splits).

@@

\begin{wrapfigure}{r}{.6\textwidth}
  %\vspace{-12ex}
  \begin{center}
    \includegraphics[width=0.6\textwidth, trim=0cm 0cm 0cm 0cm, clip=True]{figures/meg.pdf}
  \end{center}
  %\vspace{-5ex}
  \caption{Multiple models (color-coded) are compared on their ability to
  reliably predict single-trial MEG signals evoked by words. Left. Average
  improvement of correlation coefficient $\Delta R$ for each of the four
  features (rows). Error bars indicate standard error of the mean (SEM) across
  subjects. Right. Average $\Delta R$ across time for each subject (dots). Top
  horizontal lines indicate when B2B significantly outperforms other methods
  (red) and vice versa. \label{fig:meg_results}
}
\end{wrapfigure}

As expected, the Backward model reveals a similar decoding time course for Word
Length and Word Frequency, even though these features are known to specifically
influence early and late MEG responses respectively \citep{kutas2011thirty}. In
addition, the same decoding time course was observed for the dummy variable.
These results illustrate that backward modeling cannot be used to estimate the
causal contribution of collinear features.

@@

### 2020-03-04 paper/paper.tex
clean up

\documentclass{article}

\usepackage{amsmath, amsfonts, microtype, xcolor, tikz, graphicx, hyperref, amsthm}
\usepackage[ruled, linesnumbered]{algorithm2e}
\usepackage{times,wrapfig}
\usepackage{authblk}
\usepackage{apacite}
\usepackage{natbib}
\usepackage{biblatex}

@@

\title{Discriminating causes \\from multivariate observations:\\the back-to-back regression}

\author[1, 2]{Jean-Rémi King}
\author[2]{Fran\c{c}ois Charton}
\author[2]{David Lopez-Paz}
\author[2]{Maxime Oquab}
\affil[1]{Laboratoire des systèmes perceptifs, PSL University, CNRS}
\affil[2]{Facebook AI}
\date{}

% \author{%
%   Jean-Rémi King\\
%   CNRS\\
%   \texttt{jeanremi.king@gmail.com} \\
%   \And
%   Fran\c{c}ois Charton\\
%   Facebook AI\\
%   \texttt{fcharton@fb.com}\\
%   \And
%   David Lopez-Paz\\
%   Facebook AI\\
%   \texttt{dlp@fb.com}
%   \And
%   Maxime Oquab\\
%   Facebook AI\\
%   \texttt{qas@fb.com}
% }

\begin{document}

\maketitle

\begin{abstract}
Identifying causes solely from observations can be particularly challenging when
i) potential factors are difficult to manipulate individually and ii)
observations are complex and multi-dimensional. To address this issue, we
introduce ``Back-to-Back'' regression (B2B), a method designed to efficiently
measure, from a set of colinear factors, those that most plausibly account for
multidimensional observations. After proving the consistency of B2B and its
links to other linear approaches, we show that our method outperforms
least-squares regression and cross-decomposition techniques (e.g. canonical
correlation analysis and partial least squares) on causal identification.
Finally, we apply B2B to neuroimaging recordings of 102 subjects reading word
sequences. The results show that the early and late brain representations,
caused by low- and high-level word features respectively, are more reliably
detected with B2B than with other standard techniques.

\end{abstract}

@@

\include{fig_problem.tex}

We consider the measurement of multivariate signal $Y \in \mathbb{R}^{n \times
d_y}$, generated from a set of putative causes $X \in \mathbb{R}^{n \times
d_x}$, via some unknown linear apparatus $F \in \mathbb{R}^{d_x \times d_y}$.
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $E
\in \mathbb{D}^{d_x \times d_x}$, we denote by $XE$ the causal factors of $Y$.
%
In summary, the problem can be formalized as:
%
\begin{equation}
    y_i = (x_i E + n_i) F
    \label{eq:model}
\end{equation}
%
where $i$ is a given sample, and $n_i$ is a sample-specific noise drawen from a
centered distribution.
%
While the triplet of variables $X$ and $N$ are independent, we allow each of
them to have any form of covariance.
% what is a general covariance matrix?
%
In practice, we observe $n$ samples $(X, Y)$ from the model.
%
This problem space, along with the sizes of all variables involved, is
illustrated in Figure~\ref{fig:b2b}.
%
Given the model in Equation~\eqref{eq:model}, \textbf{the goal} of Back-to-Back
Regression (B2B) is to estimate the matrix of causal influences $E$.

@@

     Consider the B2B model from Equation $Y = (XE + N)F$, $N$ centred and full
     rank noise.
     %
     Let $Img(M)$ refers to the image of the matrix $M$. If $F$ and $X$ are
     full-rank on the $Img(E)$, then, the solution of B2B, $\hat H$, will
     minimize
     %
     $\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$and satisfy $E\hat H = \hat H$
\end{theorem}
%
\begin{proof}
  See Appendix \ref{appendix:theorem_proof}.
\end{proof}

@@

We perform two sets of experiments to evaluate B2B: one on controlled synthetic
data, and a second one on a real, large-scale magneto-encephalography (MEG)
dataset.
%
We use scikit-learn's PLS and RidgeCV \citep{sklearn} as well as pyrcca's
regularized canonical component analysis (RegCCA, \citep{bilenko2016pyrcca})
objects to compare B2B against the standard baselines.

@@

% I find a bit surprising to mention cauaslity literare in ML at the end. I would have expected this in the introduction to convince reviewer that know your stuff.

\section{Conclusion}
\input{conclusion}

@@

\bibliographystyle{iclr2020_conference}
\bibliography{paper}

% \input{supplementary.tex}

%\section{Appendices}
%
%This should hold part of the tests, explanations on simulation, detailed results and stuff on meg data

\input{appendix}

>>>>

\documentclass[preprint,12pt,3p]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\usepackage{amsmath, amsfonts, microtype, xcolor, tikz, graphicx, amsthm}
\usepackage[ruled, linesnumbered]{algorithm2e}
\usepackage{times,wrapfig}
\usepackage{cleveref}

@@

\journal{Neuroimage}

% \bibliographystyle{elsarticle-num}
\bibliographystyle{plain}

\begin{document}
\begin{frontmatter}

\title{Discriminating the Linear Influences of Collinear Factors \\from Multivariate Observations: the Back-to-Back Regression}

\author[1,2]{Jean-Rémi King\corref{email}}
\cortext[email]{corresponding author: jeanremi@fb.com}

\author[2]{Fran\c{c}ois Charton}
\author[2]{David Lopez-Paz}
\author[2]{Maxime Oquab}
\address[1]{Laboratoire des systèmes perceptifs, PSL University, CNRS}
\address[2]{Facebook AI}

\begin{abstract}
Identifying causes solely from observations can be particularly challenging when
i) potential factors are difficult to manipulate individually and ii)
observations are multi-dimensional. To address this issue, we
introduce ``Back-to-Back'' regression (B2B), a linear method designed to efficiently
measure, from a set of collinear factors, those that most plausibly account for
multidimensional observations. First, we prove the consistency of B2B, its links to
other linear approaches, and show how it provides a robust, unbiased and
interpretable scalar estimate for each factor.
Second, we use a variety of simulated data to show that B2B outperforms
least-squares regression and cross-decomposition techniques (e.g. canonical
correlation analysis and partial least squares) on causal identification when
both the factors and the observations are collinear.
Finally, we apply B2B to magneto-encephalography of 102 subjects
recorded during a reading task to test whether our method appropriately disentangles
the respective contribution of word length and word frequency - two collinear
factors known to cause early and late brain responses respectively. The results
show that these two factors are better disentangled with B2B than with
other standard techniques. We finally discuss the limits of B2B.
\end{abstract}

\begin{keyword}
  Cross-Decomposition \sep
  Feature Discovery \sep
  Magnetoencephalography \sep
  Decoding \sep
  Encoding \sep
  Reading \sep
  N400
\end{keyword}

\end{frontmatter}

\linenumbers

@@

\begin{figure}[t!]
  \include{fig_problem.tex}
  \label{fig:b2b}
\end{figure}

We consider the measurement of multivariate signal $Y \in \mathbb{R}^{n \times
d_y}$ (the dependent variables), generated from a set of putative causes $X \in \mathbb{R}^{n \times
d_x}$ (the independent variables), via some unknown linear apparatus $F \in \mathbb{R}^{d_x \times d_y}$.
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $E
\in \mathbb{D}^{d_x \times d_x}$, we denote by $XE$ the causal factors of $Y$.
%
In summary, the problem can be formalized as:
%
\begin{equation}
    y_i = (x_i E + n_i) F
    \label{eq:model}
\end{equation}
%
where $i$ is a given sample, and $n_i$ is a sample-specific noise drawn from a
centered distribution.
%
While the triplet of variables $X$ and $N$ are independent, we allow each of
them to have any form of covariance.
% what is a general covariance matrix?
%
In practice, we observe $n$ samples $(X, Y)$ from the model.
%
This problem space, along with the sizes of all variables involved, is
illustrated in Figure~\ref{fig:b2b}.
%
Given the model in Equation~\cref{eq:model}, the goal of Back-to-Back
Regression (B2B) is to estimate the matrix of $E$, i.e. to identify the factors
that most reliably account for the multivariate observations.

@@

     Consider the B2B model from Equation $Y = (XE + N)F$, $N$ centered and full
     rank noise.
     %
     Let $Img(M)$ refers to the image of the matrix $M$. If $F$ and $X$ are
     full-rank on the $Img(E)$, then, the solution of B2B, $\hat H$, will
     minimize
     %
     $\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$and satisfy $E\hat H = \hat H$
\end{theorem}
%
\begin{proof}
  See Appendix \ref{appendix:theorem_proof}.
\end{proof}

@@

We perform two sets of experiments to evaluate B2B: one on controlled synthetic
data, and a second one on a real, large-scale magneto-encephalography (MEG)
dataset.
%
We use scikit-learn's PLS and RidgeCV \citep{sklearn} as well as Pyrcca's
regularized canonical component analysis (RegCCA, \citep{bilenko2016pyrcca})
objects to compare B2B against the standard baselines.

@@

% I find a bit surprising to mention causality literature in ML at the end. I
% would have expected this in the introduction to convince reviewer that know
% your stuff.

\section{Conclusion}
In this work, we proposed Back-to-Back (B2B) regression, a linear method to
disentangle confounded factors from multidimensional observations.
%
B2B repeatedly performs two successive multidimensional regressions on
independent subsets of the data: the first regression is applied on the output
domain (as in backward decoding), whereas the second regression is applied on
the input domain (as in forward encoding).
%
We provided a theoretical guarantee about the consistency of B2B, and compared
it to several baselines in controlled synthetic experiments.
%
We also applied B2B to a recent brain imaging dataset, analyzing the timing of
brain responses and their connection to word features.
%
We obtained results consistent with prior work in neuroscience literature,
confirming the reliability of B2B for real data analysis.
%

@@

\bibliography{paper}

\section{Appendices}
\input{appendix}

### 2020-03-04 paper/synthetic.tex
clean up

We employ $5$-fold cross-validation to select the optimal number of components
for CCA and PLS. Regressions were $\ell2$-regularized with a $\lambda$ regularization
parameters fitted with the efficient leave-one-out procedure implemented in
scikit-learn RidgeCV \citep{sklearn}.

@@

We show in Appendix~\ref{appendix:feature_importance} pseudo-code to assess feature importance for our algorithm as well as baselines. For the Backward Model, feature importance cannot be assessed as the $X$ colinearity is never taken into account.

>>>>

We employ five-fold cross-validation to select the optimal number of components
for CCA and PLS. Regressions were $\ell2$-regularized with a $\lambda$ regularization
parameters fitted with the efficient leave-one-out procedure implemented in
scikit-learn RidgeCV \citep{sklearn}.

@@

We show in Appendix~\ref{appendix:feature_importance} pseudo-code to assess feature importance for our algorithm as well as baselines. For the Backward Model, feature importance cannot be assessed as the $X$ collinearity is never taken into account.
